---
title: "BranchGLM"
output: github_document
---

# Overview

**BranchGLM** is a package for fitting GLMs and performing branch and bound 
variable selection for GLMs.

# How to install

**BranchGLM** can be installed using the `install.packages()` function

```{r, eval = FALSE}

install.packages("BranchGLM")

```

It can also be installed via the `install_github()` function from the 
**devtools** package.

```{r, eval = FALSE}

devtools::install_github("JacobSeedorff21/BranchGLM")

```

# Usage
## Fitting GLMs
### Linear regression

**BranchGLM** can fit large linear regression models very quickly, 
next is a comparison of runtimes with the built-in `lm()` function.

```{r linear, warning = FALSE, message = FALSE, fig.path="README-"}

library(BranchGLM)
library(microbenchmark)
library(ggplot2)
set.seed(99601)

NormalSimul <- function(n, d, Bprob = .5){
  
  x <- MASS::mvrnorm(n, mu = rep(1, d), Sigma = diag(.5, nrow = d, ncol = d) + 
                 matrix(.5, ncol = d, nrow = d))
  
  beta <- rnorm(d + 1, mean = 1, sd = 1) 
  
  beta[sample(2:length(beta), floor((length(beta) - 1) * Bprob))] = 0
  
  y <- x %*% beta[-1] + beta[1] + rnorm(n, sd = 3)
  
  df <- cbind(y, x) |> 
    as.data.frame()
  
  df$y <- df$V1
  
  df$V1 <- NULL
  
  df
}
### Big simulation

df <- NormalSimul(10000, 250)

Times <- microbenchmark("BranchGLM" = {BranchGLM(y ~ ., data = df, 
                                                        family = "gaussian",
                                                   link = "identity")},
                        "Parallel BranchGLM" = {BranchGLM(y ~ ., data = df, 
                                                        family = "gaussian",
                                                   link = "identity",
                                                   parallel = TRUE)},
                        "lm" = {lm(y ~ ., data = df)},
                        times = 100)

autoplot(Times, log = FALSE)

```

### Logistic regression
**BranchGLM** can also fit large logistic regression models very quickly, 
next is a comparison of runtimes with the built-in `glm()` function. 

```{r logistic, warning = FALSE, message = FALSE, fig.path="README-"}

set.seed(78771)

LogisticSimul <- function(n, d, Bprob = .5, sd = 1){
  
  x <- MASS::mvrnorm(n, mu = rep(1, d), Sigma = diag(.5, nrow = d, ncol = d) + 
                 matrix(.5, ncol = d, nrow = d))
  
  beta <- rnorm(d + 1, mean = 0, sd = sd) 
  
  beta[sample(2:length(beta), floor((length(beta) - 1) * Bprob))] = 0
  
  p <- 1/(1 + exp(-x %*% beta[-1] - beta[1]))
  
  y <- rbinom(n, 1, p)
  
  df <- cbind(y, x) |> 
    as.data.frame()
  df
}

### Big simulation

df <- LogisticSimul(10000, 100)

Times <- microbenchmark("BFGS" = {BranchGLM(y ~ ., data = df, family = "binomial",
                                                   link = "logit", method = "BFGS")}, 
                        "L-BFGS" = {BranchGLM(y ~ ., data = df, family = "binomial",
                                                   link = "logit", method = "LBFGS")},
                        "Fisher" = {BranchGLM(y ~ ., data = df, family = "binomial",
                                                   link = "logit", method = "Fisher")},
                        "Parallel BFGS" = {BranchGLM(y ~ ., data = df, family = "binomial",
                                                   link = "logit", method = "BFGS",
                                                   parallel = TRUE)}, 
                        "Parallel L-BFGS" = {BranchGLM(y ~ ., data = df, 
                                                       family = "binomial",
                                                   link = "logit", method = "LBFGS",
                                                   parallel = TRUE)},
                        "Parallel Fisher" = {BranchGLM(y ~ ., data = df, 
                                                        family = "binomial",
                                                   link = "logit", method = "Fisher",
                                                   parallel = TRUE)},
                        "glm" = {glm(y ~ ., data = df, family = "binomial")},
                        times = 100)

autoplot(Times, log = FALSE)

```


## Variable selection

**BranchGLM** can also perform best subset selection very quickly, here is a 
comparison of runtimes with the `bestglm()` function from the **bestglm** package.

```{r, warning = FALSE, message = FALSE}

library(bestglm)
set.seed(33391)

df <- LogisticSimul(1000, 15, .5, sd = .5)

## Times
### Timing branch and bound
system.time(BranchVS <- VariableSelection(y ~ ., data = df, 
                                      family = "binomial", link = "logit",
                  type = "branch and bound", showprogress = FALSE,
                  parallel = FALSE, nthreads = 8, method = "Fisher", 
                  bestmodels = 10))

res <- summary(BranchVS)

Xy <- cbind(df[,-1], df[,1])

### Timing exhaustive search
system.time(BestVS <- bestglm(Xy, family = binomial(), IC = "AIC", TopModels = 10))

```

### Checking results

```{r}

## Results
### Checking if both methods give same results
all(names(coef(BranchVS$finalmodel)) == names(coef(BestVS$BestModel)))

```

The branch and bound method can be many times faster than an exhaustive search and 
is still guaranteed to find the optimal model.

## Parallel computation

Parallel computation can be used to greatly speed up the branch and bound 
algorithm, especially when the number of variables is large.

### Non-parallel time

```{r}

set.seed(871980)

df <- LogisticSimul(1000, 40, .5, sd = .5)

system.time(BranchVS <- VariableSelection(y ~ ., data = df, 
                                      family = "binomial", link = "logit",
                  type = "switch branch and bound", showprogress = FALSE,
                  parallel = FALSE, nthreads = 8, method = "Fisher"))
```

### Parallel time

```{r}
system.time(ParBranchVS <- VariableSelection(y ~ ., data = df, 
                                      family = "binomial", link = "logit",
                  type = "switch branch and bound", showprogress = FALSE,
                  parallel = TRUE, nthreads = 12, method = "Fisher"))

```

### Checking results

```{r}

all(names(coef(BranchVS$finalmodel)) == names(coef(ParBranchVS$finalmodel)))

```

They both give the same results
